### 2021-09-13 thesis/tex/ICintro.tex
update

% \documentclass{thesis}
% \usepackage[margin=0.5in]{geometry}
% \usepackage{graphicx, slashed, siunitx, subcaption}
% \usepackage[utf8]{inputenc}
% \usepackage{xr-hyper}
% \input{commands}
% \begin{document}
\section{Neutrino detection}
We always observe neutrinos indirectly through their associated charged lepton. Regardless of the type of interaction (charged current via the $W$ boson, or neutral current
via the $Z$), a charged lepton exits with altered properties. The lepton is then detected, and the properties of the neutrino involved in the 
interaction is then deduced. This deduction is obviously imperfect, and this introduces complexities that we will handle in Ch.~\ref{ch:ICmethod}.

@@

To detect the Cherenkov light, 60 Digital Optical Modules (DOMs) are placed on a long string up to \SI{17}{\metre} apart. 86 of these strings are then lowered
into \SI{2.5}{\km} deep boreholes in the ice. The holes are then sealed by refreezing the ice, resulting in a total of 5160 DOMs in a volume of approximately \SI{1}{\km^3}~\cite{weaverThesis}.
%TODO: pic of cherenkov light/dom?

The strings and DOMs are not spaced evenly, making some parts of the detector more sensitive to certain energy ranges than other.
8 strings packed more tightly than the other 78, making that part of the detector sensitive to neutrino energies down to single digit \si{\GeV}. Due to 
this part being situated deep within the ice, it is referred to DeepCore. DeepCore will be treated as a separate and independent detector from the rest, which
retains the name IceCube. A view of the current setup can be seen in Fig.~\ref{fig:array}. In this work, we consider DeepCore data between \SIrange{5.6}{56}{\GeV} and IceCube data in the range \SIrange{0.5}{10}{\TeV}.
\begin{figure}\label{fig:array}
    \centering
    \includegraphics[width=0.5\textwidth]{figures/icecube2.png}
    \caption{View of the full IceCube array}
\end{figure}

In 2017, the PINGU Letter of Intent was published~\cite{PINGUletter}. The `Precision IceCube Next Generation Upgrade' is an upgrade that will 
supplement DeepCore, i.e. boosing the capabilites of neutrino detection at the \si{\GeV} scale. As the PINGU upgrade is not yet financed nor built, we are
not able to use any data from it. However, the collaboration has released preliminary simulations which we will use to see how the upgrade might improve
IceCube and DeepCore bounds. The PINGU simulations have the same structure as the DeepCore data, so any analysis referring to DeepCore will
also apply to PINGU except where noted. However, we treat the PINGU detector as independent of the DeepCore experiment.

\begin{figure}[t]\label{fig:flux_aeff}
    \includegraphics[width=0.5\textwidth]{figures/flux.pdf}\hspace{-1cm}
    \includegraphics[width=0.5\textwidth]{figures/aeff.pdf}
    \vspace{-2.5cm}\caption{\emph{Left panel:} Interpolated South Pole atmospheric flux with data from~\cite{hondaArticle}.
    \emph{Right panel:} Interpolated IceCube effective area with data from~\cite{ICaeff}.}
\end{figure}

\subsection{Atmospheric Neutrino Flux}
Atmospheric neutrinos originates from cosmic rays composed of protons interacting with nuclei in the atmosphere.
These interactions ultimately produces pions, which decay as 
\begin{align}\label{eq:pion}
    &\pi^+ \to \mu^+ + \nm\,, \quad \pi^- \to \mu^- + \anm \nonumber \\
    &\pi^+ \to e^+ + \ne\,,\,\, \quad \pi^- \to e^- + \ane\,.
\end{align}
In the muonic decay channel, muons are emitted which will be detected by the IceCube detector. A part of the 
uncorrelated systematic error comes from this \emph{muon background}, i.e. events misclassified as muons from 
$\nm$ interactions rather than from pion decay. Moreover, the atmospheric flux is often associated with a large error.
In this work, we will use a flux normalization error of 24\%, and a zenith slope error of 4\%~\cite{hondapaper}.

The flux is provided in \cite{hondaData,hondaArticle}, and a selection is shown in Table~\ref{table:flux}
The flux data is binned in $\ztrue$. The fluxes are averaged over azimuthal direction and over solar minimum/maximum. 
The units of the fluxes are given as \si{\per\GeV \per\metre\squared \per\second \per\steradian} and are omitted
from the table for clarity. 
We note that the fluxes for $\nt$ and $\nu_{\bar{\tau}}$ are missing. Kaons can decay into neutral pions, which in turn 
can produce $\nt$, but this branching ratio is extremely small. Thus, we never have to use probabilities on the form 
$P_{\tau \beta}$, since we have no incoming atmospheric $\nu_\tau$ flux. 
Interpolating the data yields makes us capable of returning all four necessary fluxes for a given true energy and true zenith.
The result is shown in Fig.~\ref{fig:flux_aeff}.

\begin{table}[h]\label{table:flux}
    \begin{center}
        \begin{tabular}{lcccccc}
            \hline \hline
            $\Etrue$ [\si{\GeV}] &$\phi_\mu$ &$\phi_{\bar{\mu}}$ &$\phi_e$ &$\phi_{\bar{e}}$ & $\ztrue$\\\\
            \hline
            27825 &  \SI{6.06e-12}{} &  \SI{3.17e-12}{} &  \SI{1.56e-13}{} &  \SI{1.04e-13}{} &   [-0.2, -0.1] \\
            247707 &  \SI{5.94e-16}{} &  \SI{2.92e-16}{} &  \SI{1.36e-17}{} &  \SI{8.12e-18}{} &   [-0.7, -0.6] \\
                22 &  \SI{3.33e-02}{} &  \SI{2.78e-02}{} &  \SI{9.57e-03}{} & \SI{7.15e-03}{} &   [-0.3, -0.2] \\
            432876 &  \SI{5.19e-17}{} &  \SI{2.32e-17}{} &  \SI{1.46e-18}{} & \SI{9.83e-19}{} &   [-1.1, -1.0] \\
            64280 &  \SI{1.58e-13}{} &  \SI{8.10e-14}{} &  \SI{3.49e-15}{} &  \SI{2.21e-15}{} &   [-0.4, -0.3] \\
            \hline \hline
        \end{tabular}
    \end{center}
    \caption{A selection of processed atmospheric South Pole fluxes from~\cite{hondaData} by Honda et al.~\cite{hondaArticle}.}
\end{table}

\subsection{Event Reconstruction}
After an event has occurred, the IceCube algorithms process the data coming from the detector to \emph{reconstruct} the event. This means that, given the parameters recorded by the detector, what are their "true" values?
We are interested in two variables: the energy and the direction. Each event is tagged with a probable energy and zenith angle, called the recostructed parameters $\Ereco$ and $\zreco$, which are the parameters according to the DOMs.
The collaboration then uses numerous sophisticated methods to backtrack the reconstructed parameters to the true parameters. So a charged lepton hits the DOMs, and we ultimately end up with the associated neutrino's true and reconstructed energy and zenith angle. The reconstructed parameters are what we are using to analyze the data (because this is what the detector actually sees), while the true parameters are used in the 
determination of that neutrino's `actual' flux and cross-section (because this is what Nature sees).

How do we then translate between the reconstructed and true parameters? In this work, we are using two different methods, which are based on the form of data available to us. They will be outlined in Sec.\ref{ch:ICmethod} and Sec.~\ref{ch:DCmethod}.

>>>>

% \documentclass{thesis}
% \usepackage[margin=0.5in]{geometry}
% \usepackage{graphicx, slashed, siunitx, subcaption}
% \usepackage[utf8]{inputenc}
% \usepackage{xr-hyper}
% \input{commands}
% \begin{document}
\section{Neutrino detection}
We always observe neutrinos indirectly through their associated charged lepton. 
Regardless of the type of interaction (charged current via the $W$ boson, or neutral current
via the $Z$), a charged lepton exits with altered properties. The lepton is then detected, and the properties of the neutrino involved in the 
interaction is then deduced. This deduction will be imperfect, and this introduces complexities that we will handle in Ch.~\ref{ch:ICmethod}.

@@

To detect the Cherenkov light, 60 Digital Optical Modules (DOMs) are placed on a long string up to \SI{17}{\metre} apart. 86 of these strings are then lowered
into \SI{2.5}{\km} deep boreholes in the ice. The holes are then sealed by refreezing the ice, resulting in a total of 5160 DOMs in a volume of approximately \SI{1}{\km^3}~\cite{weaverThesis}.

The strings and DOMs are not spaced evenly, making some parts of the detector more sensitive to certain energy ranges than other.
8 strings packed more tightly than the other 78, making that part of the detector sensitive to neutrino energies down to single digit \si{\GeV}. Due to 
this part being situated deep within the ice, it is referred to DeepCore. DeepCore will be treated as a separate and independent detector from the rest, which
retains the name IceCube. A view of the current setup can be seen in Fig.~\ref{fig:array}. In this work, we consider DeepCore data between \SIrange{5.6}{56}{\GeV} and IceCube data in the range \SIrange{0.5}{10}{\TeV}.
\begin{figure}\label{fig:array}
    \centering
    \includegraphics[width=0.5\textwidth]{figures/icecube2.png}
    \caption{View of the full IceCube array, with the Eiffel tower for scale.}
\end{figure}

In 2017, the PINGU Letter of Intent was published~\cite{PINGUletter}. The `Precision IceCube Next Generation Upgrade' is an upgrade that will 
supplement DeepCore, i.e. boosing the capabilites of neutrino detection at the \si{\GeV} scale. As the PINGU upgrade is not yet financed nor built, we are
not able to use any data from it. However, the collaboration has released preliminary simulations which we will use to see how the upgrade might improve
results from IceCube and DeepCore. The PINGU simulations have the same structure as the DeepCore data, so our analysis referring to DeepCore will
also apply to PINGU except where noted. However, we treat the PINGU detector as independent of the other experiments.

\begin{figure}[t]\label{fig:flux}
    \includegraphics[width=0.7\textwidth]{figures/flux.pdf}
    \caption{Interpolated South Pole atmospheric $\nu_\mu$ flux with data from~\cite{hondaArticle}, displaying a near-neglible 
    zenith angle dependance, but a strong energy dependence. The energy dependence can be parametrized to be approximated by $E^{-2.7}$.
\end{figure}

\subsection{Atmospheric Neutrino Flux}
Atmospheric neutrinos originates from cosmic rays composed of high-energy protons interacting with nuclei in the atmosphere.
These interactions ultimately produces pions, which decay as 
\begin{align}\label{eq:pion}
    &\pi^+ \to \mu^+ + \nm\,, \quad \pi^- \to \mu^- + \anm \nonumber \\
    &\pi^+ \to e^+ + \ne\,,\,\, \quad \pi^- \to e^- + \ane\,.
\end{align}
In the muonic decay channel, muons are emitted which will be detected by the IceCube detector.
The detector's experience an uncorrelated systematic error, which comes from this \emph{muon background}, 
i.e. events misclassified as muons from 
$\nm$ interactions rather than from pion decay. For the DeepCore analysis, we will have to consider this background when calculating the events.
For IceCube events, we will scan a higher energy range where the muon background can be neglected. For the PINGU events, the IceCube detector is 
expected to be able to act as a veto for this background. Thus, the error introduced from the muon background is expected by the collaboration 
to be neglible~\cite{PINGUletter}.
Moreover, the atmospheric flux itself is often associated with a large error.
In this work, we will use a flux normalization error of 24\%, and a zenith slope error of 4\%~\cite{hondapaper}.

The flux is provided in \cite{hondaData,hondaArticle}, and a selection is shown in Table~\ref{table:flux}.
The flux data is binned in $\cos(\theta_z)$. The fluxes are averaged over azimuthal direction and over solar minimum/maximum. 
The units of the fluxes are given as \si{\per\GeV \per\metre\squared \per\second \per\steradian} and are omitted
from the table for clarity. 
We note that the fluxes for $\nt$ and $\nu_{\bar{\tau}}$ are missing. Kaons can decay into neutral pions, which in turn 
can produce $\nt$, but this branching ratio is extremely small. Thus, we never have to use probabilities on the form 
$P_{\tau \beta}$, since we have no starting atmospheric $\nu_\tau$ flux. 
Interpolating the data yields makes us capable of approximating all four necessary fluxes for a given true energy and true zenith.
The result of $\phi_\mu$ is shown in Fig.~\ref{fig:flux}. We see that the flux is almost entirely zenith independent. This is due to the 
very small cross-section of the neutrinos. For particles interacting through the electromagnetic interaction, we would expect to see a 
large difference between the flux of particles traversing through the whole diameter of the Earth, compared to particles that only penetrate a 
small rim of the crust. Here, however, the flux at the South Pole is almost the same! This does not mean that the oscillation probability will be 
equal. First, the difference in the density profiles introduces wildly different matter effects. 
Secondly, the zenith angle determines the baseline. Completely down-going neutrinos, with $\cos(\theta_z) = 0$ will be produced at an average distance of \SI{20}{\km}
overhead, travel down to the icy sheet of Antarctica, and then penetrate the ice down \SI{2}{\km} to the DOMs. 
The energy dependence of the flux is approximately $E^{-2.7}$, so our expected event count at higher energies will suffer from this strong decline in flux.

\begin{table}[h]\label{table:flux}
    \begin{center}
        \begin{tabular}{lcccccc}
            \hline \hline
            $E$ [\si{\GeV}] &$\phi_\mu$ &$\phi_{\bar{\mu}}$ &$\phi_e$ &$\phi_{\bar{e}}$ & $\cos(\theta_z)$\\\\
            \hline
            27825 &  \SI{6.06e-12}{} &  \SI{3.17e-12}{} &  \SI{1.56e-13}{} &  \SI{1.04e-13}{} &   [-0.2, -0.1] \\
            247707 &  \SI{5.94e-16}{} &  \SI{2.92e-16}{} &  \SI{1.36e-17}{} &  \SI{8.12e-18}{} &   [-0.7, -0.6] \\
                22 &  \SI{3.33e-02}{} &  \SI{2.78e-02}{} &  \SI{9.57e-03}{} & \SI{7.15e-03}{} &   [-0.3, -0.2] \\
            432876 &  \SI{5.19e-17}{} &  \SI{2.32e-17}{} &  \SI{1.46e-18}{} & \SI{9.83e-19}{} &   [-1.1, -1.0] \\
            64280 &  \SI{1.58e-13}{} &  \SI{8.10e-14}{} &  \SI{3.49e-15}{} &  \SI{2.21e-15}{} &   [-0.4, -0.3] \\
            \hline \hline
        \end{tabular}
    \end{center}
    \caption{A selection of processed atmospheric South Pole fluxes from~\cite{hondaData} by Honda et al.~\cite{hondaArticle}.}
\end{table}

\subsection{Event Reconstruction}
After an event has occurred, the IceCube algorithms process the data coming from the detector to \emph{reconstruct} the event. 
This means that, given the parameters recorded by the detector, what are their "true" values?
We are interested in two variables: the energy and the direction. Each event is tagged with a probable energy and zenith angle, 
called the recostructed parameters $\Ereco$ and $\zreco$, which are the parameters according to the DOMs.
The collaboration then uses numerous sophisticated methods to backtrack the reconstructed parameters to the true parameters. 
So a charged lepton hits the DOMs, and we ultimately end up with the associated neutrino's true energy and zenith angle, $\Etrue$ and $\ztrue$. 
The reconstructed parameters are what we are using to analyze the data (because this is what the detector actually sees), while the true parameters are used in the 
determination of that neutrino's `actual' flux and cross-section (because this is what Nature sees).

How do we then translate between the reconstructed and true parameters? 
In this work, we are using two different methods, which are based on the form of data available to us. 
They will be outlined in Sec.~\ref{ch:ICmethod} and Sec.~\ref{ch:DCmethod}.

### 2021-09-13 thesis/tex/ICtheory.tex
update

\begin{table}[h]\label{table:aeff}
    \centering
    \begin{tabular}{lrrrrr}
        \hline \hline
        $\Etrue_{min}$ [\si{\GeV}] &     $\Etrue_{max}$ [\si{\GeV}]&   $\ztrue_{min}$ &   $\ztrue_{max}$ &     $\Aeff$ [\si{\metre\squared}] \\
        \hline
             251 &      316 &  -0.92 &  -0.91 &   0.0174 \\
          794300 &  1000000 &  -0.80 &  -0.79 &  69.3600 \\
            3981 &     5012 &  -0.78 &  -0.77 &   3.1490 \\
            1585 &     1995 &  -0.07 &  -0.06 &   0.4659 \\
            398 &      501 &  -0.73 &  -0.72 &   0.0555 \\
        \hline \hline
        \end{tabular}
    \caption{IceCube-86 effective area from~\cite{ICaeff}}
\end{table}

Here, $\Aeff$ has been averaged over $\Aeff_\mu$ and $\Aeff_{\bar{\mu}}$.
Just as with the fluxes, we interpolate this in $\Etrue$, $\ztrue$ and show the result %TODO aeff interpolation

>>>>

\begin{table}[h]\label{table:aeff}
    \centering
    \begin{tabular}{lrrrrr}
        \hline \hline
        $\Etrue_{min}$ [\si{\GeV}] &     $\Etrue_{max}$ [\si{\GeV}]&   $\ztrue_{min}$ &   $\ztrue_{max}$ &     $\Aeff$ [\si{\metre\squared}] \\
        \hline
             251 &      316 &  -0.92 &  -0.91 &   0.0174 \\
          794300 &  1000000 &  -0.80 &  -0.79 &  69.3600 \\
            3981 &     5012 &  -0.78 &  -0.77 &   3.1490 \\
            1585 &     1995 &  -0.07 &  -0.06 &   0.4659 \\
            398 &      501 &  -0.73 &  -0.72 &   0.0555 \\
        \hline \hline
        \end{tabular}
    \caption{IceCube-86 effective area from~\cite{ICaeff}.}
\end{table}

Here, $\Aeff$ has been averaged over $\Aeff_\mu$ and $\Aeff_{\bar{\mu}}$.
Just as with the fluxes, we interpolate this in $\Etrue$, $\ztrue$ and show the result in Fig.~\ref{fig:aeff}
Since the IceCube array is slightly rectangular, the zenith angle affects the cross-sectional area of which the array the neutrinos are exposed to.
While the flux was almost flat in $\ztrue$, the introduction of the zenith dependent $\Aeff$ will make the result slightly more zenith dependent than 
the flux itself. Energy-wise, the situation here is completely reverse. Increasing with energy, the effective area of the detector
approaches its geometrical area of \SI{1e6}{\metre^2}, but is still only in the single digit range at \si{\TeV} energies.

\begin{figure}[t]\label{fig:aeff}
    \includegraphics[width=0.7\textwidth]{figures/aeff.pdf}
    \caption{Interpolated IceCube effective area with data from~\cite{ICaeff}.}
\end{figure}

### 2021-09-13 thesis/tex/oscillations.tex
update

Now, solving the Schrödinger equation with the Hamiltonian from Eq.~\ref{eq:H_3gen} with the matter potential from the PREM,
we obtain the nine combinations of probabilities through the Earth diameter. We note that due to our assumption of CP-invariance
(and thus, T-invariance), the probabilities $P_{\a\b}$ and $P_{\b\a}$ are identically equivalent. The result for \si{\GeV} neutrinos 
is shown in Fig.~\ref{fig:oscillations}.

@@

Now we need to incorporate the \emph{zenith angle}, defined as the angle between the neutrino direction of travel and south.
This way, neutrinos traveling through the entire diameter of the Earth up through the South Pole are defined as `up-going', while
neutrinos that start at the South Pole are `down-going'. We will mostly work with the quantity $\cos{(\theta_z)}$.
Since we are interested in the matter effects, we reserve our IceCube study to up-going neutrinos, i.e. neutrinos with
zenith angle $-1<= \cos{(\theta_z)}<=0$. We now supplement our probability grid from Fig.~\ref{fig:oscillations} with the zenith dimension,
allowing us to fully see the Earth matter effect on the oscillations in full. This is shown in Fig.~\ref{fig:oscillograms}.

>>>>

Now, solving the Schrödinger equation with the Hamiltonian from Eq.~\ref{eq:H_3gen} with the matter potential from the PREM,
we obtain the nine combinations of probabilities through the Earth diameter. We note that due to our assumption of CP-invariance
(and thus, T-invariance), the probabilities $P_{\a\b}$ and $P_{\b\a}$ are equivalent. The result for \si{\GeV} neutrinos 
is shown in Fig.~\ref{fig:oscillations}.

@@

Now we need to incorporate the \emph{zenith angle}, here defined as the angle between the neutrino direction of travel and south.
This way, neutrinos traveling through the entire diameter of the Earth up through the South Pole are defined as `up-going', while
neutrinos that start at the South Pole are `down-going'. We will mostly work with the quantity $\cos{(\theta_z)}$.
Since we are interested in the matter effects, some studies are only looking at up-going neutrinos, i.e. neutrinos with
zenith angle $-1 \le \cos{(\theta_z)} \le 0$. However, while down-going neutrinos don't experience matter oscillations (and hardly even vacuum
oscillations due to the short baseline), they can be included to reduce systematic errors.
We now supplement our probability grid from Fig.~\ref{fig:oscillations} with the zenith dimension,
allowing us to fully see the Earth matter effect on the oscillations in full. This is shown in Fig.~\ref{fig:oscillograms}.
